Bayesian Rate Estimation

### 2012-03-27 many.tex
Edits to many.tex.

We do not know a priori which of the events are foreground and which
are background.  For each event, we introduce a flag, $s_i$, which is
either 0 or 1, indicating a background or foreground event,
respectively.  These ``state'' flags are additional parameters in our
model.  We can marginalize over our uncertainty in the state of any
given event by summing all posteriors over $s_i = \{0,1\}$.  Given an
identification of each event as foreground or background, and the
rates of each, the likelihood of our data is
\begin{multline}
  p\left( d | \left\{ f(x), b(x), \left\{ s_i \right\} \right\}
  \right) = \\ \left[ \prod_{\left\{ i | s_i = 1 \right\}} f\left( x_i
    \right) \right]  \left[ \prod_{\left\{ j | s_j = 0 \right\}}
    b\left( x_j \right) \right] \\
  \times \exp\left[ - F\left( \infty \right) \right] 
  \exp\left[ -B\left( \infty \right) \right].
\end{multline}
Written in terms of the rate parameters, this becomes
\begin{multline}
  p\left( d | \left\{ f(x), b(x), \left\{ s_i \right\} \right\}
  \right) = \\ R_f^{N_f} \left[ \prod_{\left\{ i | s_i = 1 \right\}}
    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
  \times R_b^{N_b} \left[ \prod_{\left\{ j | s_j = 0 \right\}}
    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right],
\end{multline}
where $N_f$ is the number of the $s_i$ that are 1 (i.e.\ the number of
assumed foreground events), and $N_b$ is the number of the $s_i$ that
are 0 (i.e.\ the assumed number of background events).  Note that $N_f
+ N_b = N$, as each event is considered either foreground or
background in our model.

@@

A choice of the foreground and background rates implies a prior on the
state flags.  For fixed foreground and background rates, the
probability that any particular event is foreground is given by 
\begin{equation}
  p\left(s_i = 1 | R_f, R_b \right) = \frac{R_f}{R_f + R_b}.
\end{equation}

The complete prior is then
\begin{multline}
  p\left( R_f, R_b, \left\{ s_i \right\} \right) = \\ p\left( \left\{ s_i
  \right\} | R_f, R_b \right) p\left( R_f, R_b \right) = \\ \frac{R_f^{N_f} R_b^{N_b}}{\left( R_f
    + R_b\right)^N} p\left( R_f, R_b \right).
\end{multline}

@@

\section{Likelihood for Inhomogeneous Poisson Processes}

>>>>

We do not know a priori which of the events are foreground and which
are background.  For each event, we introduce a flag, $f_i$, which is
either 0 or 1, indicating a background or foreground event,
respectively.  These ``state'' flags are additional parameters in our
model.  We can marginalize over our uncertainty in the state of any
given event by summing all posteriors over $f_i = \{0,1\}$.  Given an
identification of each event as foreground or background, and the
rates of each, the likelihood%
\footnote{See \S \ref{sec:likelihood-derivation} for a derivation of
  Eq.~\eqref{eq:likelihood}.} %
of our data is
\begin{multline}
  \label{eq:likelihood}
  p\left( d, \left\{ f_i \right\} | f(x), b(x)
  \right) = \\ \left[ \prod_{\left\{ i | f_i = 1 \right\}} f\left( x_i
    \right) \right]  \left[ \prod_{\left\{ j | f_j = 0 \right\}}
    b\left( x_j \right) \right] \\
  \times \exp\left[ - F\left( \infty \right) \right] 
  \exp\left[ -B\left( \infty \right) \right].
\end{multline}
Written in terms of the rate parameters, this becomes
\begin{multline}
  p\left( d, \left\{ f_i \right\} | f(x), b(x)\right) = \\ R_f^{N_f}
  \left[ \prod_{\left\{ i | f_i = 1 \right\}}
    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
  \times R_b^{N_b} \left[ \prod_{\left\{ j | f_j = 0 \right\}}
    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right],
\end{multline}
where $N_f$ is the number of the $f_i$ that are 1 (i.e.\ the number of
assumed foreground events), and $N_b$ is the number of the $f_i$ that
are 0 (i.e.\ the assumed number of background events).  Note that $N_f
+ N_b = N$, as each event is considered either foreground or
background in our model.

@@

\subsection{Posteriors}

Putting everything together, the posterior probability of each event's
state and the overall rate is given by
\begin{multline}
  p(R_f, R_b, \left\{ f_i \right\} | d) \propto R_f^{N_f}
  \left[ \prod_{\left\{ i | f_i = 1 \right\}}
    \hat{f}\left( x_i \right) \right] \exp\left[ - R_f \right] \\
  \times R_b^{N_b} \left[ \prod_{\left\{ j | f_j = 0 \right\}}
    \hat{b}\left( x_j \right) \right] \exp\left[ - R_b \right] p(R_f, R_b).
\end{multline}
If we marginalize over all the flags, $f_i$, then we obtain
\begin{multline}
  p(R_f, R_b | d) \propto \prod_{i} \left[ R_f \hat{f}\left( x_i
    \right) + R_b \hat{b}\left( x_i \right) \right] \\ \times
  \exp\left[ - \left( R_f + R_b \right) \right] p\left( R_f, R_b
  \right).
\end{multline}
If, instead, we marginalize over all but one flag, $f_k$, we obtain
that the posterior ratio of signal to noise probability for event $k$
given the rates $R_f$ and $R_b$ is 
\begin{equation}
  \frac{p(R_f, R_b, f_k = 1 | d)}{p(R_f, R_b, f_k = 0 | d)} =
  \frac{R_f \hat{f}\left( x_k \right)}{R_b \hat{b}\left( x_k \right)}. 
\end{equation}

@@

\section{Likelihood for Inhomogeneous Poisson Processes}
\label{sec:likelihood-derivation}
